### 2015-02-04 intro.tex
Initial commit of Abstract Data Type Inference.

\section{Introduction}
\label{sec:intro}

TODO WRITE THE STORY