### 2015-02-04 related.tex

\section{Related Work}
\label{sec:related}

TODO COMPARE WITH OTHER REFINEMENT CHECKING APPROACHES AND FIGURE OUT HOW THIS
MIGHT BE RELATED TO OTHER INFERENCE ALGORITHMS