### 2015-02-27 algebra.tex
Section titles.

\section{Algebraic Properties of ADTs}
\label{sec:algebra}

TODO DEFINE EMBEDDING AND PROVE OUR ADTS ARE CLOSED

\begin{theorem}
  “Our ADTs are closed under embedding.”
\end{theorem}