### 2015-02-28 related.tex

\section{Related Work}
\label{sec:related}

TODO COMPARE WITH OTHER REFINEMENT CHECKING APPROACHES AND FIGURE OUT HOW THIS
MIGHT BE RELATED TO OTHER INFERENCE ALGORITHMS

The idea of decomposing ADT violations into a finite set of patterns has been
proposed for atomic collections~\cite{conf/concur/HenzingerSV13,
conf/popl/DoddsHK15, conf/popl/BouajjaniEEH15}, and generalized to ADTs which
can be expressed with a particular form of recursive
definitions~\cite{journals/arxiv/BouajjaniEEH15}. These works suggest patterns
which directly recognize violations across the entire set of \emph{concurrent}
executions, effectively reducing observational refinement to the verification
of classical temporal-logic properties. Some of these patterns are expressible
only as infinite, though \emph{regular}, languages. On the contrary, in this
work we suggest patterns expressed as \emph{finite} languages, which recognize
violations in ADT \emph{kernels} — i.e.,~the sequential executions of atomic
objects, or the “concurrency-aware” executions of “concurrency-aware”
specifications~\cite{conf/podc/HemedR14}. Though by themselves these patterns
are incomplete in recognizing violations across the entire set of concurrent
executions, they are complete when considered as negative conditions on
concurrent executions’ linearization orders, effectively reducing verification
of a single concurrent execution to satisfiability
checking~\cite{conf/pldi/EmmiEH15}.