### 2015-03-22 formula.tex
Formula examples.

\section{Computing Symbolic Representations}
\label{sec:formula}

TODO MOTIVATE THIS SECTION

Let $h = \tup{O,<,c,f,m,r}$ be a history with operations $O = \set{ o_1, \ldots,
o_n }$. Without loss of generality, we assume the match targets $\set{o_1,
\ldots, o_k}$ of $h$ are indexed consecutively from $1$ to $k$, for some $k \le
n$. For each $o \in O$, we define the macro $\textsc{Embed}_o(x,Y,z)$:
\begin{align*}
  & \big(c(o) \implies {\sf c}(x)\big) \land {\sf f}(x) = f(o) \land {\sf r}(x) = r(o) \\
  & \land \big(o \in \dom(m)
    \implies \lnot{\sf um}(x) \land {\sf m}(x) = z\big) \\
  & \land \big(o \not\in \dom(m) \implies {\sf um}(x)\big)
    \land \bigwedge_{y \in Y} x < y
\end{align*}
capturing the correspondence between the operation $o$ and the logical variable
$x$ representing $o$. The variables $Y$ and $z$ represent the operations
ordered after $x$, and the operation which $x$ matches. The macro
$\textsc{Identical}(x,y)$:
\begin{align*}
  & \big({\sf c}(x) \Leftrightarrow {\sf c}(y)\big)
    \land {\sf f}(x) = {\sf f}(y) 
    \land \big({\sf r}(x) \Leftrightarrow {\sf r}(y)\big) \\
  & \land \big({\sf um}(x) \Leftrightarrow {\sf um}(y)\big)
    \land {\sf m}(x) = {\sf m}(y)
\end{align*}
captures whether the operations bound to $x$ and $y$ are identical.
To express the constraints among the matches of embedded operations, we define
the macro $\textsc{Match}(Y,z)$:
\begin{align*}
  \forall x.
  {\sf m}(x) = z \implies
  {\sf r}(x) \lor \bigvee_{y \in Y} \textsc{Identical}(x,y)
\end{align*}
which requires any operation which matches $z$ to be either read-only, or
identical to some operation in $Y$, which represents the operations of $h$
which match $z$. Finally, we express the embedding of $h$ with the macro
$\textsc{Embed}_h$:
\begin{align*}
  \exists x_1, \ldots, x_n.
  \bigwedge_{i=1}^{n} \textsc{Embed}_{o_i}(x_i,Y_i,z_i)
  \land \bigwedge_{i=1}^{k} \textsc{Match}(W_i, x_i)
\end{align*}
where $Y_i = \set{ x_j : o_i < o_j }$ are the variables corresponding to
operations ordered after $o_i$, and $z_i$ is the variable corresponding to
$m(o_i)$, if defined, and $W_i = \set{ x_j : m(o_j) = o_i }$ are the variables
corresponding to operations matching $o_i$.

\begin{example}
  \label{ex:formulas}

  Consider again the histories of Example~\ref{ex:patterns} which generate the
  complement of the atomic register ADT. The $\textsc{Embed}$ formula for the
  first history,
\begin{verbatim}
  [1:X] read => 1 (RO)  #
\end{verbatim}
  after simplifications, like replacing $\mathrm{true} \implies p$ with $p$, is
  \begin{align*}
    \exists x_1.\ {\sf c}(x_1) \land {\sf f}(x_1) = \mathrm{read} \land {\sf um}(x_1).
  \end{align*}
  The $\textsc{Embed}$ formula for second history,
\begin{verbatim}
  [1:1] write(1)            #
  [2:2] read => empty (RO)    #    
\end{verbatim}
  is similarly given by
  \begin{align*}
    & \exists x_1, x_2.\ x_1 < x_2 \\
    & \quad \land {\sf c}(x_1) \land {\sf f}(x_1) = \mathrm{write}
      \land \lnot{\sf um}(x_1) \land {\sf m}(x_1) = x_1 \\
    & \quad \land {\sf c}(x_2) \land {\sf f}(x_2) = \mathrm{read}
      \land \lnot{\sf um}(x_2) \land {\sf m}(x_2) = x_2 \\
    & \quad \land \big(\forall x.\ {\sf m}(x) = x_1 \implies {\sf r}(x) \lor \textsc{Identical}(x,x_1)\big) \\
    & \quad \land \big(\forall x.\ {\sf m}(x) = x_2 \implies {\sf r}(x) \lor \textsc{Identical}(x,x_2)\big).
  \end{align*}
  The formulas for the other histories are similarly obtained.

\end{example}

\begin{lemma}

  $h_1 \models \textsc{Embed}_{h_2}$ if{f} $h_1 \succeq h_2$.

\end{lemma}

We obtain a formula representing an ADT $A$ by taking the conjunction of the
negative embedding formulas $\set{ \lnot\textsc{Embed}_h : h \in H }$ from any
set $H$ that generates the complement of $A$.

\begin{lemma}

  The formula $\bigwedge_{h \in H} \neg \textsc{Embed}_h$ represents $A$,
  for any set $H$ which generates the compliment of $A$.

\end{lemma}

\begin{example}

  The conjunction of negations of the $\textsc{Embed}$ formula for the
  histories of Example~\ref{ex:patterns}, which are partially written in
  Example~\ref{ex:formulas}, represents the atomic register ADT.

\end{example}

For normal and predictable ADTs, such sets are computable, and so our inference
problem is also computable.

\begin{theorem}

  The symbolic ADT inference problem is computable for normal and predictable ADTs.

\end{theorem}