### 2015-03-22 discussion.tex
More on nature.

\section{Discussion of Limitations}
\label{sec:discussion}

NOTE LIMITATIONS: PRIORITY QUEUE, SEMAPHORE, RENDEZVOUS SYNCHRONIZER,
EXCHANGER, BARRIER, …