### 2015-03-22 matching.tex
More motivation paragraphs.

\section{Computing Matching Schemes}
\label{sec:matching}

Our solution to the ADT inference problem relies on identifying faithful
matching schemes for ADT implementations. Though Section~\ref{sec:nature} shows
such matching schemes exist for naturally-occurring ADTs, the next natural
question with regard to automation is whether these matching schemes can be
generated automatically.

In this section we demonstrate that matching schemes can in fact be generated
systematically with minimal manual specification by reduction to logical
satisfiability. Essentially, we formulate matching schemes as uninterpreted
functions in satisfiability queries constrained by the requirement that they
\emph{normalize} an implementation’s executions, i.e.,~generate histories which
are closed under the operation removals of Section~\ref{sec:patterns}. This
requirement is given with respect to an enumeration of execution pairs in which
the first is admitted by the given implementation, and the second is not, yet
it is a projection of the first’s operations. Consequently, this prohibits any
normalizing matching scheme from considering the projected operations a match.
In what follows, we suppose the read-only component $R$ of matching schemes
$\tup{M,R}$ is given, and focus on the generation of the per-execution matching
functions $M$.

The required manual specification includes identifying a class of executions to
which an implementation should be subjected, and a set of predicates that are
required in the logic of matching schemes. Formally, a \emph{language} $L =
\tup{E,P,Q}$ is a set $E$ of executions, along with finite sequences $P = P_1
P_2 \ldots$ and $Q = Q_1 Q_2 \ldots$ of binary and ternary predicates
$P_i(e,o_1)$ and $Q_i(e,o_1,o_2)$ ranging over the executions of $E$ and their
operations. When $P$ is a $k$-length sequence of $n$-ary predicates, we write
$P(x_1, \ldots, x_n)$ to denote the valuation sequence
\begin{align*}
  P_1(x_1, \ldots, x_n) \ldots P_k(x_1, \ldots, x_n).
\end{align*}
Given a language $L = \tup{E,P,Q}$, we say a matching scheme $M$ is
\emph{simple} when there exists an $n$-ary Boolean function $G$, for $n =
2\cdot|P|+|Q|$, such that for each execution $e \in E$
\begin{itemize}

  \item $G(P(e,o_1),P(e,o_2),Q(e,o_1,o_2))$ is satisfied for at most one
  operation $o_1$, for any operation $o_2$,

  \item $M(e)(o_2)$ is undefined unless there exists an operation $o_1$
  for which $G(P(e,o_1),P(e,o_2),Q(e,o_1,o_2))$, and

  \item $M(e)(o_2) = o_1$ if{f} $G(P(e,o_1),P(e,o_2),P(e,o_1,o_2))$,

\end{itemize}
where $o_1$ and $o_2$ range over the operations of $e$.

\begin{example}
  \label{ex:register-language}

  We say an execution of read and write methods \emph{writes unique values} if
  the argument value to each write operation is unique. Consider the language
  whose executions write unique values, with the following predicates:
  \begin{align*}
    & \mathsf{w}(e,o) & \text{$o$ is a {\sf write} operation in $e$, and}, \\
    & \mathsf{v}(e,o_1,o_2) & \text{$o_1$ and $o_2$ read/write the same values.}
  \end{align*}
  We define the function $G(x_\mathsf{w}, y_\mathsf{w}, z_\mathsf{v})$ over
  valuations to the predicates above to be satisfied if and only if:
  \begin{align*}
    x_\mathsf{w} \land z_\mathsf{v}
  \end{align*}
  Intuitively, this defines a simple matching scheme for which write
  operations match themselves, and read operations match the write which wrote
  the value read. In the case such a write exists, it is unique in any
  execution which writes unique values. The match is otherwise undefined. This
  scheme is exactly that which is specified in Example~\ref{ex:matching} of
  Section~\ref{sec:histories}.

\end{example}

An implementation $\mathcal{I}$ \emph{adheres} to a language $L = \tup{E,P,Q}$
if $\mathcal{I} \subseteq E$. A match scheme $M$ \emph{normalizes} an
implementation $\mathcal{I}$ when $M$ is faithful to $\mathcal{I}$ and
$H(\mathcal{I},M)$ is normal.

\begin{definition}

  The \emph{matching scheme inference problem} is to compute a simple matching
  scheme $M$ which normalizes a given implementation $\mathcal{I}$ adhering to
  a given language $L$.

\end{definition}

We automate the computation of matching schemes by considering the function
${\sf g}$ underlying a simple matching scheme as an uninterpreted function in a
logical satisfiability problem which uniquely determines the match of each
operation, and constrain ${\sf g}$ to normalize the executions of an
implementation. To this end, we fix an implementation $\mathcal{I}$ and a
language $L = \tup{E,P,Q}$ to which $\mathcal{I}$ adheres, then consider any
enumeration $F$ of execution pairs $\tup{e,e'} \in E^2$ such that
\begin{itemize}

  \item $e \in \mathcal{I}$ and $e' \not\in \mathcal{I}$, and

  \item $e'$ is obtained by deleting operations of $e$.

\end{itemize}
Any such pair of executions can be used to rule out several possibilities,
for instance, that the operations removed from $e$ to obtain $e'$:
\begin{itemize}

  \item do not constitute a match in $e$,

  \item are not all duplicate operations,

  \item are not all read-only operations,

  \item do not constitute multiple matches in $e$,

\end{itemize}
and so on. Ruling out these possibilities is sound since, for example, a
normalized scheme $M$ could not consider those operations a match: otherwise
the history abstraction $H(\mathcal{I},M)$ which includes $H(e,M)$ must also
include $H(e',M)$, being normal, and in particular closed under match removal.
Such an $M$ would thus not be faithful. For simplicity, in what follows we
consider ruling out only the first possibility: that the operations removed
from $e$ are not a match. In principle, the approach extends to rule out all
possibilities.

In what follows, we denote the operations of an execution $e$ by $O_e$.
To consider whether a given pair $o_1, o_2$ of operations of an execution $e$
is a match according to the simple matching scheme based on ${\sf g}$, for
the given language $L$, we define the macro $\textsc{IsMatch}_{e,o_1,o_2}$:
\begin{align*}
  \mathsf{g}\big( P(e,o_1), P(e,o_2), Q(e,o_1,o_2) \big).
\end{align*}
To enforce that each operation of an execution $e$ will have a unique match,
we define the macro $\textsc{UniqueMatch}_e$:
\begin{align*}
  \bigwedge_{o_1,o_2 \in O_e} \Big(
    \textsc{IsMatch}_{e,o_2,o_2} \implies
    \bigwedge_{o_1' \neq o_1} \lnot \textsc{IsMatch}_{e,o_1',o_2}
  \Big).
\end{align*}
Then a given set $O \subseteq O_e$ constitutes a match according to
${\sf g}$ when all operations $o_2 \in O$ target some operation $o_1 \in O$,
and no other operation $o_2 \in O \setminus O_e$ does. We express this with the
macro $\textsc{EntireMatch}_{e,O}$:
\begin{align*}
  \bigvee_{o_1 \in O}
  \Big(
  \bigwedge_{o_2 \in O} \textsc{IsMatch}_{e,o_1,o_2}
  \land \bigwedge_{o_2 \in O_e \setminus O} \lnot \textsc{IsMatch}_{e,o_1,o_2}
  \Big)
\end{align*}
Finally, given a pair $\tup{e,e'} \in F$, we prohibit the operations $O_e
\setminus O_{e'}$ from constituting a match according to ${\sf g}$, since if
${\sf g}$ normalized $\mathcal{I}$, and $e \in \mathcal{I}$, then $e'$ should
also be in $\mathcal{I}$. We express this exclusion for all pairs of $F$
with the macro $\textsc{Normalizes}_{F}$:
\begin{align*}
  \bigwedge_{\tup{e,e'}\in F} \Big(
    \textsc{UniqueMatch}_{e}
    \land \lnot \textsc{EntireMatch}_{e,(O_e \setminus O_{e'})}
  \Big).
\end{align*}
We thus check satisfiability for the conjunction of non-matches $O_e \setminus
O_{e'}$ over all pairs $\tup{e,e'} \in F$.

\begin{example}

  Consider again the language $L_\mathrm{reg}$ of read and write methods of
  Example~\ref{ex:register-language} with predicates ${\sf w}(e,o)$ and ${\sf
  v}(e,o_1,o_2)$ whose executions write unique values. The following table
  lists all possible predicate valuations $\tup{x_{\sf w}, y_{\sf w}, z_{\sf
  v}}$, and for each valuation a valid execution which excludes the positive
  valuation of ${\sf g}(x_{\sf w}, y_{\sf w}, z_{\sf v})$ in the satisfaction
  of $\textsc{Normalize}_F$, in the case such an execution exists.
  \begin{center}
    \begin{tabular}{ccc|ll}
      $x_\mathsf{w}$ & $y_\mathsf{w}$ & $z_\mathsf{v}$ & counterexample & reason \\
      \hline
      $0$ & $0$ & $0$ & ${\rm w}(1) \ {\rm r}(1) \ {\rm r}(1) \ {\rm w}(2) \ {\rm r}(2)$ & not unique \\
      $0$ & $0$ & $1$ & ${\rm w}(1) \ {\rm r}(1) \ {\rm r}(1)$ & not unique \\
      $0$ & $1$ & $0$ & ${\rm w}(1) \ \underline{{\rm r}(1) \ {\rm w}(2)} \ {\rm r}(2)$ & $\to_\mathrm{m}$ \\
      $0$ & $1$ & $1$ & $\underline{{\rm w}(1) \ {\rm r}(1)} \ {\rm r}(1)$ & $\to_\mathrm{m}$ \\
      $1$ & $0$ & $0$ & ${\rm w}(1) \ {\rm w}(2) \ {\rm w}(3) \ {\rm r}(3)$ & not unique \\
      $1$ & $0$ & $1$ & & \\
      $1$ & $1$ & $0$ & ${\rm w}(1) \ {\rm w}(2) \ {\rm w}(3)$ & not unique \\
      $1$ & $1$ & $1$ & & \\
    \end{tabular}
  \end{center}
  For example, the first valuation $000$ must be excluded since it allows the
  read ${\rm r}(2)$ to match both reads ${\rm r}(1)$, violating
  $\textsc{UniqueMatch}$. The second valuation $001$ must also be excluded,
  since it allows either read ${\rm r}(1)$ to match both itself and the other
  read ${\rm r}(1)$. The third valuation $010$ must be excluded since it allows
  the write ${\rm w}(2)$ to match the read ${\rm r}(1)$, in which case removing
  the match $\set{ {\rm r}(1), {\rm w}(2) }$ results in the invalid execution
  ${\rm w}(1)\ {\rm r}(2)$. Reasoning follows similarly for the other rows.
  In this way, any enumeration $F$ which includes the above executions
  will exclude all valuations except for $101$ and $111$, ultimately resulting
  in a satisfiable $\textsc{Normalize}_F$ in which ${\sf g}(x_{\sf w}, y_{\sf
  w}, z_{\sf v})$ is the same Boolean function $x_{\sf w} \land z_{\sf v}$
  given in Example~\ref{ex:register-language}.

\end{example}

On the one hand, checking satisfiability of $\textsc{Normalizes}_F$ can be used
to conclude the impossibility of a good matching scheme — at least for the
given language.

\begin{lemma}

  If $\textsc{Normalizes}_F$ is unsatisfiable, then there exists no simple
  matching scheme that normalizes $\mathcal{I}$ for the language $L$.

\end{lemma}

In this case, the reason for unsatisfiability can be used as a counterexample
to refine the given language, for example, by adding additional predicates.

\begin{example}

  Consider again the language $L_\mathrm{reg}$ of read and write methods of
  Example~\ref{ex:register-language}, yet this time without the predicate ${\sf
  v}(e,o_1,o_2)$. The following table lists all possible predicate valuations
  $\tup{x_{\sf w}, y_{\sf w}}$, and for each valuation a valid execution which
  excludes the positive valuation of ${\sf g}(x_{\sf w}, y_{\sf w})$ in the
  satisfaction of $\textsc{Normalize}_F$.
  \begin{center}
    \begin{tabular}{cc|ll}
      $x_{\sf w}$ & $y_{\sf w}$ & counterexample & reason \\
      \hline
      0 & 0 & ${\rm w}(1)\ {\rm r}(1)\ {\rm r}(1)$ & not unique \\
      0 & 1 & $\underline{{\rm w}(1)\ {\rm r}(1)}\ {\rm r}(1)$ & $\to_\mathrm{m}$ \\
      1 & 0 & ${\rm w}(1)\ {\rm w}(2)\ {\rm r}(2)$ & not unique \\
      1 & 1 & ${\rm w}(1)\ {\rm w}(2)$ & not unique \\
    \end{tabular}
  \end{center}
  Thus any enumeration $F$ including the executions above excludes every
  possible valuation of $\tup{x_{\sf w}, y_{\sf w}}$, resulting in an
  unsatisfiable $\textsc{Normalize}_F$.

\end{example}

On the other hand, the satisfiability of $\textsc{Normalizes}_F$ does not
necessarily lead to a unique matching scheme, since $\textsc{Normalize}_F$ can
have multiple satisfying assignments. Furthermore, $\textsc{Normalizes}_F$ does
not necessarily normalize $\mathcal{I}$. For one reason, $F$ may not be a
complete set of examples of executions and invalid projections. Second, our
simple characterization of $\textsc{Normalize}_F$ does not rule out the other
reasons for a given example $\tup{e,e'} \in F$ to be an invalid projection,
e.g.,~that the removed operations do not constitute \emph{multiple} matches.
However, we believe that checking satisfiability of $\textsc{Normalize}_F$ is
useful nonetheless: at the very least, satisfying assignments can be used as
\emph{assistance} in constructing normalizing matching schemes.