### 2015-03-23 discussion.tex
Discussion section.

\section{Discussion}
\label{sec:discussion}

While the theoretical foundation proposed in this work covers the majority of
naturally-occurring ADTs, it is worth mentioning a few conceptual limitations
and possible solutions to overcome them.

First, while our theoretical foundation does cover “concurrency-aware” ADT
specifications~\cite{conf/podc/HemedR14}, i.e.,~non-atomic ADTs like the
\emph{rendezvous synchronizer}, \emph{barrier}, and \emph{exchanger}, whose
kernels have bounded width greater than $1$, our current implementation only
handles atomic ADTs. In principle, this limitation is not fundamental. The key
difference is in the enumeration of kernel histories. For a given ADT,
determining whether a given history is in the kernel or its complement reduces
to checking whether there exists an execution of which the history is an
abstraction. For atomic ADTs, only a single sequential execution need be
examined, in which no two operations overlap. For non-atomic ADTs, every
possible interleaving of the internal actions of operations need be examined.
While we expect this calculation to remain feasible given that ADT complements
are normally represented with histories with few operations, we do expect it to
incur a noticeable cost.

Second, ADTs whose specifications rely on relations (besides equality) on
method argument and return values cannot be expressed with our notion of
histories based on matching schemes. A notable example of such an ADT is the
\emph{priority queue}, whose \emph{dequeue} operations return the
smallest/largest \emph{enqueue}d value which has not yet been dequeued. It
appears that such ADTs are rare in practice. Overcoming this limitation would
require a refinement to matching schemes which can track additional relations
among data values.

Finally, in some cases it is unclear how to express matching schemes
deterministically. For example, each \emph{wait} operation of \emph{semaphore}
ADT should naturally match the \emph{notify} operation which enabled it.
However, it is not possible to determine this based on operation labels alone,
and it is not clear whether all implementations effectively keep track of this
correspondence. Furthermore, choosing matches arbitrarily would be unsound, and
result in inferring ADTs with artificial constraints, e.g.,~that the notify and
wait operations behave according to a FIFO discipline, each wait matching the
oldest unmatched notify. Overcoming this limitation may require introducing a
notion of nondeterminism into matching schemes.