### 2015-06-26 motivation.tex
Adding motivation section.

\section{Motivation}
\label{sec:motivation}

TODO DISCUSS SCALABLE VERIFICATION, MODULAR REASONING, INTERFACE SPECIFICATIONS,
DERIVING SPECIFICATIONS, CONCURRENCY, CONCURRENT ADTS, SYMBOLIC ADT SPECIFICATIONS

TODO CLIENT-SIDE REASONING

\begin{figure}
  \begin{minipage}{0.49\linewidth}
    \begin{verbatim}
assume w < x
( (q.add(<1,w>); q.add(<1,x>))
||(q.add(<2,w>); q.add(<2,x>)))
i, y := q.remove()
j, z := q.remove()
assert i == j ==> y < z
    \end{verbatim}
  \end{minipage}
  \hfill
  \begin{minipage}{0.49\linewidth}
    \begin{verbatim}
( (while * do q.add(<1,w++>))
||(while * do q.add(<2,x++>)))
while * do q.remove() end
i, y := q.remove()
j, z := q.remove()
assert i == j ==> y < z
    \end{verbatim}
  \end{minipage}
  \caption{A atomic queue clients with six operations (left), and with
  arbitrarily-many operations (right).}
  \label{fig:clients}
\end{figure}

\begin{align*}
  \forall w,x,y,z.\left(
    \begin{array}{l}
      {\sf add}(w) \land {\sf remove}(x) \land {\sf val}(w) = {\sf val}(x) \\
      \land\ {\sf add}(y) \land {\sf remove}(z) \land {\sf val}(y) = {\sf val}(z) \\
      \land\ {\sf before}(w,y) \implies {\sf before}(x,z)
    \end{array}
  \right)
\end{align*}

TODO IMPLEMENTATION-SIDE REASONING: EMP(6 < 4) AND TR(4 < 6)

\begin{verbatim}
    [1] add(1)        #
    [2] remove => 3     #####
    [3] remove => 1     #
    [4] add(2)            #
    [5] add(3)              #
    [6] remove => EMPTY       #
\end{verbatim}

\begin{align*}
  \forall x,y.\left(
    \begin{array}{l}
      {\sf remove}(x) \land {\sf val}(x) = {\sf empty} \\
      \land\ {\sf add}(y) \land {\sf unmatched}(y) \implies {\sf before}(x,y)
    \end{array}
  \right)
\end{align*}

TODO AUTOMATIC INFERENCE