### 2015-06-30 motivation.tex
Nearly complete draft of motivation section.

\section{Motivation}
\label{sec:motivation}

Effective scalable reasoning about nontrivial software implementations generally
requires considering each software module separately, in isolation, using
abstract specifications for other modules. When modules are objects whose
methods may be called concurrently, their behavior is typically understood in
terms of valid invocation sequences of atomic abstract data types (ADTs): any
execution of overlapping method invocations can be \emph{linearized} into an
invocation sequence admitted by the ADT~\cite{journals/toplas/HerlihyW90}. ADT
specifications thus decouple reasoning about object implementations — is each
concurrent execution linearizable? — from their clients’ invocations — does
every valid linearization preserve client invariants?

Consider the two parallel programs in Figure~\ref{fig:clients} invoking the add
and remove methods of an atomic queue implementation. The clients add increasing
integer values $w$ and $x$ tagged with the integers $\set{1,2}$ indicating on
which parallel branch each add occurs. Intuitively both programs are correct
since values with the same tag are added in increasing order, and, crucially,
the values of the queue ADT are removed in the same order in which they are
added. Thus considering the sequences of values added by all six possible
linearizations of the add operations for the first program:
\begin{align*}
  & \texttt{<1,w> <1,x> <2,w> <2,x>}, \\
  & \texttt{<1,w> <2,w> <1,x> <2,x>}, \\
  & \texttt{<1,w> <2,w> <2,x> <1,x>}, \\
  & \texttt{<2,w> <1,w> <1,x> <2,x>}, \\
  & \texttt{<2,w> <1,w> <2,x> <1,x>}, \\
  & \texttt{<2,w> <2,x> <1,w> <1,x>}
\end{align*}
we see that comparison {\tt i == j} of tags for the first two removes only holds
for the first and last linearizations, which remove $w$ then $x$ according to
the queue ADT specification, implying that {\tt y < z} holds when {\tt i == j}
holds.

\begin{figure}
  \begin{minipage}{0.49\linewidth}
    \begin{verbatim}
assume w < x
( (q.add(<1,w>); q.add(<1,x>))
||(q.add(<2,w>); q.add(<2,x>)))
i, y := q.remove()
j, z := q.remove()
assert i == j ==> y < z
    \end{verbatim}
  \end{minipage}
  \hfill
  \begin{minipage}{0.49\linewidth}
    \begin{verbatim}
( (while * do q.add(<1,w++>))
||(while * do q.add(<2,x++>)))
while * do q.remove() end
i, y := q.remove()
j, z := q.remove()
assert i == j ==> y < z
    \end{verbatim}
  \end{minipage}
  \caption{A atomic queue clients with six operations (left), and with
  arbitrarily-many operations (right).}
  \label{fig:clients}
\end{figure}

ADTs are often defined in an imperative style using simple mathematical objects.
For example, the atomic queue ADT can be specified via an abstract varaible $q$
initialized to the empty sequence $\varepsilon$; add operations update $q$ to
$q' = q \cdot x$ by appending the argument value $x$; when $q = x \cdot q'$,
remove operations update $q$ to $q'$ and return $x$; otherwise, when $q =
\varepsilon$, remove operations return a special value {\tt EMPTY}. Reasoning
using such specifications requires the explicit consideration of possible
linearizations one by one in order to determine valid return values, e.g.,~for
the remove operations of queues.

In this work we advocate \emph{symbolic} ADT specifications which facilitate the
use of efficient reasoning algorithms, e.g.,~via symbolic reasoning engines. For
simplicity we consider simple first-order languages whose variables represent
\emph{operations}, i.e.,~method invocations, with predicates indicating the
method name and argument/return values of each operation, along with the
linearization order of operations. For example, the following formula partially
specifies the atomic queue ADT: the removes matching a pair of add operations
occur in the same order as their corresponding adds:
\begin{align*}
  \forall w,x,y,z.\left(
    \begin{array}{l}
      {\sf add}(w) \land {\sf remove}(x) \land {\sf val}(w) = {\sf val}(x) \\
      \land\ {\sf add}(y) \land {\sf remove}(z) \land {\sf val}(y) = {\sf val}(z) \\
      \land\ {\sf before}(w,y) \implies {\sf before}(x,z)
    \end{array}
  \right).
\end{align*}
Such symbolic representations are more convenient than the aforementioned
imperative style of ADT specification since they allow symbolic reasoning
engines to prune redundant reasoning steps. For example, rather than enumerating
the six linearizations above of the first program of Figure~\ref{fig:clients} to
determine that the valid return values of remove operations are $w$ and $x$, a
symbolic reasoning engine could eagerly rule out all but the first and last
linearizations for which {\tt i == j} by looking at only the first two
operations. Since the number of linearizations is generally exponential in the
number of operations, such symbolic reasoning is of significant value.

TODO THE SECOND CLIENT?

Symbolic ADT specifications also facilitate reasoning about the correctness of
object \emph{implementations} by avoiding the redundant reasoning inherent in
the explicit enumeration of linearizations~\cite{conf/pldi/EmmiEH15}. For
example, consider the execution log of an erroneous queue implementation
\begin{verbatim}
    [1] add(1)        #
    [2] remove => 3     #####
    [3] remove => 1     #
    [4] add(2)            #
    [5] add(3)              #
    [6] remove => EMPTY       #
\end{verbatim}
where each operation is prefixed by a unique operation identifier; the “\#”
symbol denotes the time span of each operation, from left to right, indicating
that Operation~$2$ overlaps in time with Operations~$3$, $4$, and~$5$, and that
all other operations are ordered in time. While this execution can be identified
as a violation of the atomic queue ADT by considering every possible
linearization of Operation~$2$, this approach is intractible in general, as
the number of linearizations is exponential in the number of operations.
However, consider the following formula partially specifying the atomic
queue ADT: that the adds of values which have not yet been removed must be
ordered after empty remove operations:
\begin{align*}
  \forall x,y.\left(
    \begin{array}{l}
      {\sf remove}(x) \land {\sf val}(x) = {\sf empty} \\
      \land\ {\sf add}(y) \land {\sf unmatched}(y) \implies {\sf before}(x,y)
    \end{array}
  \right).
\end{align*}
Combined with the transitivity of the {\sf before} relation, a symbolic
reasoning engine could eagerly conclude that Operation~$4$ must be linearized
both before Operation~$6$, due to transitivity, and after, due to the previous
axiom. Such inconsistencies can be detected with great efficiency in
practice~\cite{conf/pldi/EmmiEH15}.

While (symbolic) ADT specifications are crucial to scalable reasoning about
nontrivial software implementations, writing effective speciications is often a
challenge for programmers. Towards fuller automation of scalable reasoning
techniques, we focus in this work on the automatic inference of symbolic
specifications which can be used directly in both client-side and
implementation-side reasoning of concurrent objects.