### 2015-07-06 algorithm.tex
starting top-down presentation

%!TEX root = draft.tex
\section{Solving the Symbolic ADT Inference Problem}

\begin{figure}
\begin{lstlisting}
Min := $\emptyset$
while (true) {
  $B_i$ := $ \set{ h \in B : \weight(h) = i }$;
  MinTemp := Min $\cup$ Min($B_i$)
  if ( MinTemp = Min )
    break;
  Min := MinTemp
}
return $\neg$Formula(Min);
\end{lstlisting}
\caption{The algorithm solving the symbolic ADT inference problem for a given ADT $A$. $B$ denotes the bounded complement of $\ker A$ and $\uparrow$\texttt{Min} denotes the upward closure w.r.t. $\preceq$ of a given set of histories. The procedure $\texttt{Min}(X)$ returns the $\preceq$-minimal elements of the set of histories $X$ and $\texttt{Formula}(X)$ returns a history formula which holds exactly for the histories in the upward closure of $X$.} 
\label{fig:algorithm}
\end{figure}

Figure~\ref{fig:algorithm} lists an algorithm for solving the symbolic ADT inference problem which is parametrized by 
an order relation $\preceq$ between histories such that the complement of the ADT 
is closed under $\preceq$. For a given set of elements $X$ and an order relation $\preceq$, the \emph{upward closure}
of $X$ w.r.t. $\preceq$ is the set $\uparrow X=\set{y : \exists x\in X.\ x\preceq y}$. The algorithm performs an
weight-increasing enumeration of $B$ while maintaining the
$\preceq$-minimal elements, and stops whenever each history of weight $i$ is bigger (w.r.t. $\preceq$) than some
element of the collected set of minimals.

The partial correctness of the algorithm (i.e., ignoring termination) relies on the input ADT being \emph{predictable}, i.e.,
for all $i$,
\begin{align*}
B_{i+1} \succeq \texttt{Min}(B_{\leq i}) \mbox{ implies } \texttt{Min}(B) = \texttt{Min}(B_{\leq i}),
\end{align*}
where $B_{\leq i}=\bigcup_{1\leq j\leq i}B_j$.

\begin{theorem}\label{th:corr1}
For any predictable ADT $A$, if the algorithm in Figure~\ref{fig:algorithm} terminates, then the returned formula represents $A$.
\end{theorem}
\begin{proof}
%For a set $\Sigma$, let $\neg \Sigma$ denote its complement.
Let $M$ be the value of \texttt{Min} at the end of the \texttt{while} loop.
Since the complement of $A$ is closed under $\preceq$, and $M\subseteq B\subseteq  \neg \ker A \subseteq \neg A$, we have that $\uparrow M\subseteq \neg A$. 
Since $A$ is predictable, we have that $B \subseteq\, \uparrow M$.
The formula $\varphi$ returned by the algorithm holds exactly for the set of histories $X$ 
which is the complement of $\uparrow M$. By the previous inclusions, we have that $A\subseteq X\subseteq \neg B$.
Therefore, $\varphi$ holds for all histories in $\ker A$ and doesn't hold for histories in $B$. 
\end{proof}

\begin{remark}
Note that for non-predictable ADTs, the value of $\texttt{Min}$ at the end of the \texttt{while} loop 
is an under-approximation of $\texttt{Min}(B)$. 
The resulting symbolic ADT representation is still satisfied by all histories in $\ker A$, therefore it would still be sound in only identifying
\emph{actual} violations, yet incomplete in identifying \emph{all} violations.
\end{remark}

The following result states the conditions under which the algorithm in Figure~\ref{fig:algorithm} terminates.

\begin{theorem}\label{th:corr2}
If the set of $\preceq$-minimal elements of any (potentially-infinite) set of histories is finite, then the algorithm in Figure~\ref{fig:algorithm} terminates.
\end{theorem}

The following section provides a concrete instantiation for the order relation $\preceq$ satisfying the requirements in 
Theorem~\ref{th:corr1} and~\ref{th:corr2} except for predictability. 
Although we don't provide a general result concerning predictability, we prove that it holds for many naturally-occurring ADTs (see Section~\ref{sec:nature}).