### 2015-07-06 motivation.tex
Reworked the motivation.

\section{Motivation}
\label{sec:motivation}

Effective scalable reasoning about nontrivial software implementations generally
requires considering each software module separately, in isolation, using
abstract specifications for other modules. When modules are objects whose
methods may be called concurrently, their behavior is typically understood in
terms of valid invocation sequences of atomic abstract data types (ADTs): any
execution of overlapping method invocations can be \emph{linearized} into an
invocation sequence admitted by the ADT~\cite{journals/toplas/HerlihyW90}. For
example, consider the execution history depicted in Figure~\ref{fig:clients} in
which the add operations numbered $2$ and $3$ overlap with each other, and,
respectively, with operations $1$ and $4$. Among the five possible ways to
linearize these six operations, the linearization~$1$; $3$; $2$; $4$; $5$; $6$
does correspond to a valid sequence of the atomic queue ADT, and so this history
is linearizable. ADT specifications thus decouple reasoning about object
implementations — is each concurrent execution linearizable? — from their
clients’ invocations — do all valid linearizations preserve client invariants?

\begin{figure}
  \begin{minipage}{0.43\linewidth}
    \begin{verbatim}
1: add(a)       #
2: add(b)       ###
3: add(c)         ###
4: add(d)           #
5: remove => a        #
6: remove => c          #
    \end{verbatim}
  \end{minipage}
  \hfill
  \begin{minipage}{0.55\linewidth}
    \begin{verbatim}
assume w < x
(  (q.add(<1,w>); q.add(<1,x>))
|| (q.add(<2,w>); q.add(<2,x>)) )
i, y := q.remove()
j, z := q.remove()
assert i == j ==> y < z
    \end{verbatim}
  \end{minipage}
  \caption{An execution history with six numbered operations (left),
    and a parallel program invoking six operations (right). The hash symbols
    “\#” depict the time interval spanned by each operation horizontally.}
  \label{fig:clients}
\end{figure}

Consider the parallel program in Figure~\ref{fig:clients} invoking the add and
remove methods of an atomic queue implementation, adding increasing integer
values $w$ and $x$ tagged with the integers $\set{1,2}$ indicating on which
parallel branch each add occurs. Intuitively this program is correct since
values with the same tag are added in increasing order, and, crucially, the
values of the queue ADT are removed in the same order in which they are added.
Among the six possible ways to linearize these operations, the comparison {\tt i ==
j} of tags only holds for those two which begin with {\tt q.add(<1,$w$>);
q.add(<1,$x$>)} and {\tt q.add(<2,$w$>); q.add(<2,$x$>)}. Since the queue ADT
dictates that elements are removed in the order added, we conclude that $w$ and
$x$ are removed in order when {\tt i == j}, and thus {\tt y < z} holds when {\tt
i == j} holds.

Although formal ADT specifications are indispensable for scalable program
reasoning, formal-specification writing is a burden for which few programmers
posses the required combination of expertise and willingness to overcome.
According to current practices, programmers write obviously-correct ADT
\emph{reference implementations}, e.g.,~an atomic queue whose methods are
synchronized via a global lock, and refine them with more efficient fine-grained
implementations, e.g.,~reducing synchronization overhead with specialized
hardware instructions such as atomic compare and swap (CAS).

Our goal in this work is thus the automated generation of formal ADT
specifications, derived from reference implementations, which are suitable for
automated reasoning. In particular, we aim to generate \emph{symbolic
representations} of the history linearizations which are valid, according to
ADTs given implicitly by reference implementations. We target declarative
symbolic representations rather than imperative state-based representations in
order to exploit efficient symbolic reasoning engines such as SMT solvers.
Previous work demonstrates that symbolic representations can offer orders of
magnitude improvements over imperative state-based representations by averting
the need to enumerate linearizations explicitly in order to check their
validity~\cite{conf/pldi/EmmiEH15}.

TODO ALGORITHM OVERVIEW