### 2015-07-06 motivation.tex
Abstract algorithm.

\section{Motivation \& Overview}
\label{sec:motivation}

Effective scalable reasoning about nontrivial software implementations generally
requires considering each software module separately, in isolation, using
abstract specifications for other modules. When modules are objects whose
methods may be called concurrently, their behavior is typically understood in
terms of valid invocation sequences of atomic abstract data types (ADTs): any
execution of overlapping method invocations can be \emph{linearized} into an
invocation sequence admitted by the ADT~\cite{journals/toplas/HerlihyW90}. For
example, consider the execution history depicted in Figure~\ref{fig:clients} in
which the add operations numbered $2$ and $3$ overlap with each other, and,
respectively, with operations $1$ and $4$. Among the five possible ways to
linearize these six operations, the linearization~$1$; $3$; $2$; $4$; $5$; $6$
does correspond to a valid sequence of the atomic queue ADT, and so this history
is linearizable. ADT specifications thus decouple reasoning about object
implementations — is each concurrent execution linearizable? — from their
clients’ invocations — do all valid linearizations preserve client invariants?

\begin{figure}
  \begin{minipage}{0.43\linewidth}
    \begin{verbatim}
1: add(a)       #
2: add(b)       ###
3: add(c)         ###
4: add(d)           #
5: remove => a        #
6: remove => c          #
    \end{verbatim}
  \end{minipage}
  \hfill
  \begin{minipage}{0.55\linewidth}
    \begin{verbatim}
assume w < x
(  (q.add(<1,w>); q.add(<1,x>))
|| (q.add(<2,w>); q.add(<2,x>)) )
i, y := q.remove()
j, z := q.remove()
assert i == j ==> y < z
    \end{verbatim}
  \end{minipage}
  \caption{An execution history with six numbered operations (left),
    and a parallel program invoking six operations (right). The “\#” symbols
    depict the time intervals spanned by operations horizontally.}
  \label{fig:clients}
\end{figure}

Consider the parallel program in Figure~\ref{fig:clients} invoking the add and
remove methods of an atomic queue implementation, adding increasing integer
values $w$ and $x$ tagged with the integers $\set{1,2}$ indicating on which
parallel branch each add occurs. Intuitively this program is correct since
values with the same tag are added in increasing order, and, crucially, the
values of the queue ADT are removed in the same order in which they are added.
Among the six possible ways to linearize these operations, the comparison {\tt i ==
j} of tags only holds for those two beginning with
\begin{quote}
  \verb|q.add(<1,w>); q.add(<1,x>) | and \verb| q.add(<2,w>); q.add(<2,x>)|
\end{quote}
Since the queue ADT dictates that elements are removed in the order added, we
conclude that $w$ and $x$ are removed in order when {\tt i == j}, and thus {\tt
y < z} holds when {\tt i == j} holds.

Although formal ADT specifications are indispensable for scalable program
reasoning, formal-specification writing is a burden for which few programmers
posses the required combination of expertise and willingness to overcome.
Typically programmers write simple ADT \emph{reference implementations},
e.g.,~whose methods are synchronized via a global lock, and refine them with
more efficient fine-grained implementations, e.g.,~reducing synchronization
bottlenecks using specialized hardware instructions such as atomic compare and
swap (CAS).

Our goal in this work is thus the automated generation of formal ADT
specifications, derived from reference implementations, which are suitable for
automated reasoning. In particular, we aim to generate \emph{symbolic
representations} of the valid history linearizations, according to ADTs given
implicitly by reference implementations. We target declarative symbolic
representations rather than imperative state-based representations in order to
harness efficient symbolic reasoning algorithms: rather than enumerating
linearizations explicitly, and checking their validity one by one, a symbolic
reasoning engine may simultaneously rule out many possible linearizations,
e.g.,~as in the example above. Previous work demonstrates that such symbolic
reasoning can increase reasoning efficiency by orders of
magnitude~\cite{conf/pldi/EmmiEH15}.

Our basic approach to inferring ADT specifications, as outlined by the abstract
algorithm in Figure~\ref{fig:abstract}, is to identify a finite set of
sequential execution histories which capture all of the reasons for which a
sequence could be considered invalid, according to the ADT of a given reference
implementation. These sequences thus serve as patterns indicating violations in
the sequences which contain them. Thus the linearizations which exclude all
violation patterns are considered valid.

\begin{figure}
  \begin{verbatim}
procedure INFER_ADT_FROM_REFERENCE_IMPL(I)
  PATTERNS := {}
  for each sequential history H
    if * then
      break
    else if H is EXECUTABLE with I then
      continue
    else if H is REDUNDANT with PATTERNS then
      continue
    else
      add H to PATTERNS
    end
  end
  return EXCLUSION of PATTERNS
  \end{verbatim}
  \caption{An abstract algorithm for symbolic ADT inference.}
  \label{fig:abstract}
\end{figure}

The key technical obstacle in realizing the abstract algorithm of
Figure~\ref{fig:abstract}, overcome in Section~\ref{sec:patterns}, is the
classification of violations into a finite set of patterns, rendering the
remaining invalid sequences redundant. The symbolic representation which
excludes a given set of patterns is relatively straightforward to construct, and
is given in Section~\ref{sec:formula}. Note however that the termination of this
abstract algorithm is nondeterministic, and thus the inferred ADT specification
is generally not the strongest possible in the sense that it may not exclude
certain invalid linearizations. Nevertheless, specifications inferred by this
algorithm are sound, in the sense that when modular program reasoning with
inferred specifications succeeds, correctness follows. In any case, the
specifications inferred by our refinement of this abstract algorithm in
Section~\ref{sec:algorithm} are both sound and complete for the
naturally-occurring ADTs of interest.