### 2015-07-07 patterns.tex

%!TEX root = draft.tex
\section{Computing Finite ADT Representations}
\label{sec:patterns}

In this section we demonstrate that naturally-occurring ADTs can be precisely
represented by finite sets of histories, despite the fact that these ADTs admit
infinite sets of histories. This results is nontrivial, and relies on the
identification of certain algebraic properties of the sets of histories
admitted by ADTs, more precisely, that they are
closed under the removal of certain operations. 
%and that the bounded complement of their kernel has 
%sets of histories
%%adhere to a well-founded ordering under the relation induced by such removals.
%have 
%a finite set of minimal elements with respect to the relation induced by such removals.

In addition to the relations $\to_\mathrm{o}$, $\to_\mathrm{p}$, and
$\to_\mathrm{c}$ of Section~\ref{sec:histories} under which all ADTs are
closed, the ADTs we consider in this work are also closed under additional
relations which remove read-only operations, unmatched operations, entire
matches, and duplicate operations. The relations $\to_\mathrm{r}$,
$\to_\mathrm{u}$, $\to_\mathrm{m}$ and $\to_\mathrm{d}$ relate two histories
$h_1$ and $h_2$ when $h_2$ is obtained from $h_1$ by
\begin{itemize}

  \item removing a read-only operation (r),

  \item removing an unmatched operation (u),

  \item removing a match (m), or

  \item removing a duplicate operation (d).

\end{itemize}
We say an ADT closed under $\to_\mathrm{r}$, $\to_\mathrm{u}$, $\to_\mathrm{m}$
and $\to_\mathrm{d}$ is \emph{normal}. In Section~\ref{sec:nature} we
demonstrate that naturally-occurring ADTs are normal. Defining the relation
$\succeq$ as the reflexive and transitive closure of the above relations,
\begin{align*}
  \mathord{\succeq} = (
    \to_\mathrm{o} \cup \to_\mathrm{p} \cup \to_\mathrm{c} \cup 
    \to_\mathrm{r} \cup \to_\mathrm{u} \cup \to_\mathrm{m} \cup \to_\mathrm{d}
  )^\ast
\end{align*}
closure under $\succeq$ follows immediately.

\begin{lemma}\label{lem:ADT_closure}

  Normal ADTs are closed under $\succeq$.

\end{lemma}

Besides this closure property, the $\preceq$ relation enjoys a certain notion
of well-foundedness (when restricted to a certain class of histories): every potentially 
infinite set of width-bounded histories has a finite set of minimal elements w.r.t. $\preceq$.

\begin{lemma}\label{lem:wqo}

  Sets of width-bounded histories have a finite set of $\preceq$-minimals.

\end{lemma}

For the remainder of this section, we fix an ADT $A$, and let $B$ be the bounded
complement of $\ker A$. When $\ker A$ has bounded width, so does $B$, and thus
$B$ has a finite set of $\preceq$-minimals. 
Furthermore, when $A$ normal, it is closed under
$\succeq$, and thus $B$ is closed under 
$\preceq$~\footnote{Actually, $B$ is closed under $\preceq$ when restricted to width-bounded histories, i.e., if $h_1 \preceq h_2$, $h_1 \in B$, and $h_2$ is width-bounded, then $h_2\in B$.}.
Closure under a relation satisfying Lemma~\ref{lem:wqo} implies
representation by a finite set. Formally, we say a set $X$ is \emph{finitely
representable} if there exists a finite set $Y$ and a relation $R \subseteq Y
\times X$ such that $X = \set{ x : \exists y\in Y.\ R(y,x) }$. In our case, we
obtain a finite set from which exactly the elements of $B$ are related by
$\preceq$.

\begin{lemma}

  $B$ is finitely representable if $A$ is normal.

\end{lemma}

%Computing $B$ thus amounts to computing its finite representation. In general,
%this is achieved by enumerating the elements of $B$ while maintaining the
%$\preceq$-minimal elements, and recognizing a condition under which all
%elements of $B$ are related to the current set of
%minimals~\cite{conf/lics/AbdullaCJT96, journals/tcs/FinkelS01}. In our case, we
%recognize a condition on the weight-increasing enumeration of $B$ which
%guarantees coverage by the current set of minimals. Formally, for all $i \in
%\mathbb{N}$, we define
%\begin{align*}
%  B_i &= \set{ h \in B : \weight(h) \le i } \\
%  B_i' &= \set{ h' \in B :
%    \exists h \in B_i.\ h \preceq h' \text{ and } \weight(h') \le i+1
%  }
%\end{align*}
%as, respectively, the histories of $B$ with at most $i$ matches and duplicates,
%and those derived from $B_i$ with at most $i\!+\!1$ matches and duplicates. We
%say that $A$ is \emph{predictable} if $B_i = B$ whenever $B_i' = B_{i+1}$. In
%Section~\ref{sec:nature} we demonstrate that naturally-occurring ADTs are
%predictable. Since the number of weight-$i$ histories is finite\footnote{As
%noted in Section~\ref{sec:histories}, we consider equality between histories up
%to renaming of operation identifiers.} for each $i \in \mathbb{N}$, each $B_i$
%is computable.
%
%\begin{lemma}
%
%  $B_i$ is computable, for all $i \in \mathbb{N}$, if $A$ is normal.
%
%\end{lemma}
%
%Predictability ensures that once $B_{i+1}$ does not alter the set of computed
%$\preceq$-minimals, with respect to those of $B_i$, then the minimals computed
%for $B_i$ are exactly those of $B$.
%
%\begin{lemma}
%
%  $B$ is computable if $A$ is normal and predictable.
%
%\end{lemma}
%
%Note that for non-predictable ADTs, computing $B_i$, for any $i \in
%\mathbb{N}$, under-approximates $B$. In practice, computing $B_i$ can be used
%to identify all violations which surface in histories with $i$ matches,
%overlooking violations which only surface with more than $i$ matches. The
%resulting symbolic ADT representation would still be sound in only identifying
%\emph{actual} violations, yet incomplete in identifying \emph{all} violations.

\begin{example}
  \label{ex:patterns}

  The following four histories generate the complement of the atomic
  single-value register ADT, witnessing either an unmatched read operation:
\begin{verbatim}
  [1:X] read => 1 (RO)  #
\end{verbatim}
  a read of an uninitialized register occurring after some write:
\begin{verbatim}
  [1:1] write(1)            #
  [2:2] read => - (RO)    #
\end{verbatim}
  a read which happens before its matching write operation:
\begin{verbatim}
  [1:2] read => 1 (RO)  #
  [2:2] write(1)          #
\end{verbatim}
  and a read matching a write which is not the most recent:
\begin{verbatim}
  [1:1] write(1)        #
  [2:2] write(2)          #
  [3:1] read => 1 (RO)      #
\end{verbatim}
  Every sequential history not admitted by the atomic register ADT embeds at
  least one of these four histories.

\end{example}