### 2015-10-25 draft.tex
Applied SIGPLAN formatting guidelines.

\documentclass[9pt,natbib,authoryear]{sigplanconf}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{listings}
\usepackage[ruled]{algorithm2e}

\lstset{
  basicstyle=\ttfamily,
  mathescape
}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\img}{img}
\DeclareMathOperator{\width}{width}
\DeclareMathOperator{\weight}{weight}
\DeclareMathOperator{\freq}{freq}

\newcommand{\set}[1]{{\{{#1}\}}}
\newcommand{\mset}[1]{{\{\!\{ #1 \}\!\}}}
\newcommand{\tup}[1]{{\left\langle{#1}\right\rangle}}
\renewcommand{\implies}{\Rightarrow}

\makeatletter
\renewcommand{\verbatim@font}{\normalfont\ttfamily\scriptsize}
\makeatother

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

% \conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country}
% \copyrightyear{20yy}
% \copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
% \doi{nnnnnnn.nnnnnnn}
\toappear{}

% Uncomment one of the following two, if you are not going for the
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish,
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers,
                                  % short abstracts)

% \titlebanner{banner above paper title}        % These are ignored unless
% \preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Symbolic Abstract Data Type Inference}
% \subtitle{}

\authorinfo{Michael Emmi}
           {IMDEA Software Institute, Spain}
           {michael.emmi@imdea.org}

\authorinfo{Constantin Enea}
           {Université Paris Diderot, France}
           {cenea@univ-paris-diderot.fr}

\maketitle

\begin{abstract}

  Formal specification is a vital ingredient to scalable verification of
  software systems. In the case of efficient implementations of concurrent
  objects like atomic registers, queues, and locks, symbolic formal
  representations of their abstract data types (ADTs) enable efficient modular
  reasoning, decoupling clients from implementations. Writing adequate formal
  specifications, however, is a complex task requiring rare expertise. In
  practice, programmers write reference implementations as informal
  specifications.

  In this work we demonstrate that effective symbolic ADT representations can
  be automatically generated from the executions of reference implementations.
  Our approach exploits two key features of naturally-occurring ADTs: violations
  can be decomposed into a small set of representative patterns, and these
  patterns manifest in executions with few operations. By identifying certain
  algebraic properties of naturally-occurring ADTs, and exhaustively sampling
  executions up to a small number of operations, we generate concise symbolic
  ADT representations which are complete in practice, enabling the application
  of efficient symbolic verification algorithms without the burden of manual
  specification. Furthermore, the concise ADT violation patterns we generate
  are human-readable, and can serve as useful, formal documentation.

\end{abstract}

\category{F.3.1}{Specifying and Verifying and Reasoning about Programs}{Mechanical verification}

\terms
Reliability, Verification

\keywords
Concurrency; Refinement; Linearizability

\input intro
\input overview
\input histories
\input inference
\input patterns
\input formula
\input algorithm
\input matching
\input nature
\input discussion
\input related

\appendix
\input appendix

% \acks …

% \balancecolumns

\bibliographystyle{abbrvnat}
\bibliography{violin}

% % The bibliography should be embedded for final submission.
%
% \begin{thebibliography}{}
% \softraggedright
%
% \bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
% P. Q. Smith, and X. Y. Jones. ...reference text...
%
% \end{thebibliography}

\end{document}