### 2015-10-25 inference.tex
Corrections from Reviewer B.

%!TEX root = draft.tex
\section{The Symbolic ADT Inference Problem}
\label{sec:inference}

In this section we formalize a notion of abstract data type and define the
corresponding refinement and inference problems. These are the foundational
problems addressed in this work.

A Set $K$ \emph{generates} $H$ when the closure $K^*$ of $K$ under the relation
$\mathord\to = (\to_\mathrm{o} \cup \to_\mathrm{p} \cup \to_\mathrm{c})$
is equal to $H$, i.e., $H = \set{ h : \exists h' \in K.\ h' \to^\ast h }$.
A \emph{kernel} of a set $H$ is a minimal set generating $H$. While the kernels
of arbitrary sets need not be unique, the kernels of sets which
have sequential kernels are unique. Furthermore, Section~\ref{sec:nature}
demonstrates that the histories of naturally-occurring implementations have
unique kernels, which we assume for the remainder of this work.
An \emph{abstract data type (ADT)} $A$ is the kernel of the set $H(\mathcal{I})$
of histories of some implementation $\mathcal{I}$; we say that \emph{$A$ is the
ADT of $\mathcal{I}$}.

ADTs and reference implementations serve as specifications to more efficient
implementations in the sense that they limit the set of histories that efficient
implementation may admit. This notion of refinement ensures that client program
(safety) properties which hold using the ADT or reference implementation also
hold using refined implementations~\cite{conf/popl/BouajjaniEEH15}.

\begin{definition}

  An implementation $\mathcal{I}_1$ \emph{refines} another implementation
  $\mathcal{I}_2$ when $H(\mathcal{I}_1) \subseteq H(\mathcal{I}_2)$. An
  implementation $\mathcal{I}$ refines an ADT $A$ when $H(\mathcal{I})
  \subseteq A^*$.

\end{definition}

Recent works demonstrate efficient\footnote{In time polynomial in the number of
operations, per execution.} refinement-checking
algorithms~\cite{conf/popl/BouajjaniEEH15, conf/pldi/EmmiEH15} yet rely on
hand-written symbolic ADT representations. In order to frame the problem of
computing these automatically, we fix a language for symbolic representation. A
\emph{history formula} is a first-order logic formula with
\begin{itemize}

  \item variables ranging over operation identifiers,

  \item constants from $\mathbb{M}$ for method names,

  \item function symbols ${\sf f}$ and ${\sf m}$ for labels and matching, and

  \item predicate symbols ${\sf c}$, ${\sf um}$, {\sf r}, and ${\sf <}$ for completion,
  non-matching (operations which are not in the domain of the matching function), read-only, and order.

\end{itemize}
A history formula $F$ is interpreted over a history $h$ in the natural way, by
binding variables to the operations of $h$, and binding function and predicate
symbols to their interpretations in $h$. We write $h \models F$ when $h$ is a
model of $F$, and $h \not\models F$ otherwise.

\begin{example}

  The following history formula is satisfied by histories in which no write
  operation happens between a pair of matching write and read operations:
  \begin{align*}
    \forall x_1, x_2, x_3.\
    & {\sf c}(x_1) \land {\sf f}(x_1) = \mathrm{write} \land {\sf m}(x_1) = x_1 \\
    & \land {\sf c}(x_2) \land {\sf f}(x_2) = \mathrm{write} \land {\sf m}(x_2) = x_2 \\
    & \land {\sf c}(x_3) \land {\sf f}(x_3) = \mathrm{read} \land {\sf m}(x_3) = x_1 \\
    & \land x_1 < x_2
      \implies x_3 < x_2
  \end{align*}
  This is one of several requirements of atomic single-value register ADTs.
  It is satisfied by certain linearizations of the histories $H(e_2)$ and
  $H(e_3)$ from Example~\ref{ex:histories-of-executions}, yet not $H(e_1)$.

\end{example}

The \emph{bounded complement} of a history set $H$ of width $k \in \mathbb{N} \cup
\set{\omega}$ is the set of histories of width at most $k$ which are excluded
from $H$. Let $A$ be an ADT and $B$ its bounded complement. We say that
a history formula $F$ \emph{represents} $A$ when
\begin{itemize}

  \item $h \models F$ for all $h \in A$, and

  \item $h \not\models F$ for all $h \in B$.

\end{itemize}
ADT inference is to compute a formula representing an ADT.

\begin{definition}

  The \emph{symbolic ADT inference problem} is to compute a history formula
  representing the ADT of a given implementation.

\end{definition}

Computing a history formula representing the ADT of a reference implementation
$\mathcal{I}$ thus enables efficient modular program reasoning without the
burden of writing precise formal specifications for $\mathcal{I}$.