### 2015-11-02 draft.tex
Final version candidate.

\documentclass[9pt,natbib,authoryear]{sigplanconf}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{listings}
\usepackage[ruled]{algorithm2e}

\lstset{
  basicstyle=\ttfamily,
  mathescape
}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\img}{img}
\DeclareMathOperator{\width}{width}
\DeclareMathOperator{\weight}{weight}
\DeclareMathOperator{\freq}{freq}

\newcommand{\set}[1]{{\{{#1}\}}}
\newcommand{\mset}[1]{{\{\!\{ #1 \}\!\}}}
\newcommand{\tup}[1]{{\left\langle{#1}\right\rangle}}
\renewcommand{\implies}{\Rightarrow}

\makeatletter
\renewcommand{\verbatim@font}{\normalfont\ttfamily\scriptsize}
\makeatother

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

% \conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country}
% \copyrightyear{20yy}
% \copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
% \doi{nnnnnnn.nnnnnnn}
\toappear{}

% Uncomment one of the following two, if you are not going for the
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish,
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers,
                                  % short abstracts)

% \titlebanner{banner above paper title}        % These are ignored unless
% \preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Symbolic Abstract Data Type Inference}
% \subtitle{}

\authorinfo{Michael Emmi}
           {IMDEA Software Institute, Spain}
           {michael.emmi@imdea.org}

\authorinfo{Constantin Enea}
           {Université Paris Diderot, France}
           {cenea@univ-paris-diderot.fr}

\maketitle

\begin{abstract}

  Formal specification is a vital ingredient to scalable verification of
  software systems. In the case of efficient implementations of concurrent
  objects like atomic registers, queues, and locks, symbolic formal
  representations of their abstract data types (ADTs) enable efficient modular
  reasoning, decoupling clients from implementations. Writing adequate formal
  specifications, however, is a complex task requiring rare expertise. In
  practice, programmers write reference implementations as informal
  specifications.

  In this work we demonstrate that effective symbolic ADT representations can
  be automatically generated from the executions of reference implementations.
  Our approach exploits two key features of naturally-occurring ADTs: violations
  can be decomposed into a small set of representative patterns, and these
  patterns manifest in executions with few operations. By identifying certain
  algebraic properties of naturally-occurring ADTs, and exhaustively sampling
  executions up to a small number of operations, we generate concise symbolic
  ADT representations which are complete in practice, enabling the application
  of efficient symbolic verification algorithms without the burden of manual
  specification. Furthermore, the concise ADT violation patterns we generate
  are human-readable, and can serve as useful, formal documentation.

\end{abstract}

\category{F.3.1}{Specifying and Verifying and Reasoning about Programs}{Mechanical verification}

\terms
Reliability, Verification

\keywords
Concurrency; Refinement; Linearizability

\input intro
\input overview
\input histories
\input inference
\input patterns
\input formula
\input algorithm
\input matching
\input nature
\input discussion
\input related

\appendix
\input appendix

% \acks …

% \balancecolumns

\bibliographystyle{abbrvnat}

% \bibliography{dblp,books}

\begin{thebibliography}{26}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Abdulla et~al.(1996)Abdulla, Cerans, Jonsson, and
  Tsay]{conf/lics/AbdullaCJT96}
P.~A. Abdulla, K.~Cerans, B.~Jonsson, and Y.~Tsay.
\newblock General decidability theorems for infinite-state systems.
\newblock In \emph{Proceedings, 11th Annual {IEEE} Symposium on Logic in
  Computer Science, New Brunswick, New Jersey, USA, July 27-30, 1996}, pages
  313--321. {IEEE} Computer Society, 1996.
\newblock \url{http://dx.doi.org/10.1109/LICS.1996.561359}.

\bibitem[Abdulla et~al.(2013)Abdulla, Haziza, Hol{\'{\i}}k, Jonsson, and
  Rezine]{conf/tacas/AbdullaHHJR13}
P.~A. Abdulla, F.~Haziza, L.~Hol{\'{\i}}k, B.~Jonsson, and A.~Rezine.
\newblock An integrated specification and verification technique for highly
  concurrent data structures.
\newblock In \emph{Tools and Algorithms for the Construction and Analysis of
  Systems - 19th International Conference, {TACAS} 2013, Held as Part of the
  European Joint Conferences on Theory and Practice of Software, {ETAPS} 2013,
  Rome, Italy, March 16-24, 2013. Proceedings}, volume 7795 of \emph{Lecture
  Notes in Computer Science}, pages 324--338. Springer, 2013.
\newblock \url{http://dx.doi.org/10.1007/978-3-642-36742-7_23}.

\bibitem[Amit et~al.(2007)Amit, Rinetzky, Reps, Sagiv, and
  Yahav]{conf/cav/AmitRRSY07}
D.~Amit, N.~Rinetzky, T.~W. Reps, M.~Sagiv, and E.~Yahav.
\newblock Comparison under abstraction for verifying linearizability.
\newblock In \emph{Computer Aided Verification, 19th International Conference,
  {CAV} 2007, Berlin, Germany, July 3-7, 2007, Proceedings}, volume 4590 of
  \emph{Lecture Notes in Computer Science}, pages 477--490. Springer, 2007.
\newblock \url{http://dx.doi.org/10.1007/978-3-540-73368-3_49}.

\bibitem[Angluin(1987)]{journals/iandc/Angluin87}
D.~Angluin.
\newblock Learning regular sets from queries and counterexamples.
\newblock \emph{Inf. Comput.}, 75\penalty0 (2):\penalty0 87--106, 1987.
\newblock \url{http://dx.doi.org/10.1016/0890-5401(87)90052-6}.

\bibitem[Bouajjani et~al.(2015{\natexlab{a}})Bouajjani, Emmi, Enea, and
  Hamza]{conf/icalp/BouajjaniEEH15}
A.~Bouajjani, M.~Emmi, C.~Enea, and J.~Hamza.
\newblock On reducing linearizability to state reachability.
\newblock In \emph{Automata, Languages, and Programming - 42nd International
  Colloquium, {ICALP} 2015, Kyoto, Japan, July 6-10, 2015, Proceedings, Part
  {II}}, volume 9135 of \emph{Lecture Notes in Computer Science}, pages
  95--107. Springer, 2015{\natexlab{a}}.
\newblock \url{http://dx.doi.org/10.1007/978-3-662-47666-6_8}.

\bibitem[Bouajjani et~al.(2015{\natexlab{b}})Bouajjani, Emmi, Enea, and
  Hamza]{conf/popl/BouajjaniEEH15}
A.~Bouajjani, M.~Emmi, C.~Enea, and J.~Hamza.
\newblock Tractable refinement checking for concurrent objects.
\newblock In \emph{Proceedings of the 42nd Annual {ACM} {SIGPLAN-SIGACT}
  Symposium on Principles of Programming Languages, {POPL} 2015, Mumbai, India,
  January 15-17, 2015}, pages 651--662. {ACM}, 2015{\natexlab{b}}.
\newblock \url{http://doi.acm.org/10.1145/2676726.2677002}.

\bibitem[Burckhardt et~al.(2010)Burckhardt, Dern, Musuvathi, and
  Tan]{conf/pldi/BurckhardtDMT10}
S.~Burckhardt, C.~Dern, M.~Musuvathi, and R.~Tan.
\newblock Line-up: a complete and automatic linearizability checker.
\newblock In \emph{Proceedings of the 2010 {ACM} {SIGPLAN} Conference on
  Programming Language Design and Implementation, {PLDI} 2010, Toronto,
  Ontario, Canada, June 5-10, 2010}, pages 330--340. {ACM}, 2010.
\newblock \url{http://doi.acm.org/10.1145/1806596.1806634}.

\bibitem[Burnim et~al.(2011)Burnim, Necula, and Sen]{conf/asplos/BurnimNS11}
J.~Burnim, G.~C. Necula, and K.~Sen.
\newblock Specifying and checking semantic atomicity for multithreaded
  programs.
\newblock In \emph{Proceedings of the 16th International Conference on
  Architectural Support for Programming Languages and Operating Systems,
  {ASPLOS} 2011, Newport Beach, CA, USA, March 5-11, 2011}, pages 79--90.
  {ACM}, 2011.
\newblock \url{http://doi.acm.org/10.1145/1950365.1950377}.

\bibitem[Dodds et~al.(2015)Dodds, Haas, and Kirsch]{conf/popl/DoddsHK15}
M.~Dodds, A.~Haas, and C.~M. Kirsch.
\newblock A scalable, correct time-stamped stack.
\newblock In \emph{Proceedings of the 42nd Annual {ACM} {SIGPLAN-SIGACT}
  Symposium on Principles of Programming Languages, {POPL} 2015, Mumbai, India,
  January 15-17, 2015}, pages 233--246. {ACM}, 2015.
\newblock \url{http://doi.acm.org/10.1145/2676726.2676963}.

\bibitem[Dragoi et~al.(2013)Dragoi, Gupta, and Henzinger]{conf/cav/DragoiGH13}
C.~Dragoi, A.~Gupta, and T.~A. Henzinger.
\newblock Automatic linearizability proofs of concurrent objects with
  cooperating updates.
\newblock In \emph{Computer Aided Verification - 25th International Conference,
  {CAV} 2013, Saint Petersburg, Russia, July 13-19, 2013. Proceedings}, volume
  8044 of \emph{Lecture Notes in Computer Science}, pages 174--190. Springer,
  2013.
\newblock \url{http://dx.doi.org/10.1007/978-3-642-39799-8_11}.

\bibitem[Emmi et~al.(2015)Emmi, Enea, and Hamza]{conf/pldi/EmmiEH15}
M.~Emmi, C.~Enea, and J.~Hamza.
\newblock Monitoring refinement via symbolic reasoning.
\newblock In \emph{Proceedings of the 36th {ACM} {SIGPLAN} Conference on
  Programming Language Design and Implementation, Portland, OR, USA, June
  15-17, 2015}, pages 260--269. {ACM}, 2015.
\newblock \url{http://doi.acm.org/10.1145/2737924.2737983}.

\bibitem[Finkel and Schnoebelen(2001)]{journals/tcs/FinkelS01}
A.~Finkel and P.~Schnoebelen.
\newblock Well-structured transition systems everywhere!
\newblock \emph{Theor. Comput. Sci.}, 256\penalty0 (1-2):\penalty0 63--92,
  2001.
\newblock \url{http://dx.doi.org/10.1016/S0304-3975(00)00102-X}.

\bibitem[Fishburn(1985)]{books/Fishburn85}
P.~C. Fishburn.
\newblock \emph{Interval Orders and Interval Graphs: A Study of Partially
  Ordered Sets}.
\newblock Wiley-Interscience Series in Discrete Mathematics. John Wiley \& Sons
  Inc, 1985.

\bibitem[Giannakopoulou et~al.(2005)Giannakopoulou, Pasareanu, and
  Barringer]{journals/ase/GiannakopoulouPB05}
D.~Giannakopoulou, C.~S. Pasareanu, and H.~Barringer.
\newblock Component verification with automatically generated assumptions.
\newblock \emph{Autom. Softw. Eng.}, 12\penalty0 (3):\penalty0 297--320, 2005.
\newblock \url{http://dx.doi.org/10.1007/s10515-005-2641-y}.

\bibitem[Hemed and Rinetzky(2014)]{conf/podc/HemedR14}
N.~Hemed and N.~Rinetzky.
\newblock Brief announcement: concurrency-aware linearizability.
\newblock In \emph{{ACM} Symposium on Principles of Distributed Computing,
  {PODC} '14, Paris, France, July 15-18, 2014}, pages 209--211. {ACM}, 2014.
\newblock \url{http://doi.acm.org/10.1145/2611462.2611513}.

\bibitem[Hendler et~al.(2006)Hendler, Lev, Moir, and
  Shavit]{journals/dc/HendlerLMS06}
D.~Hendler, Y.~Lev, M.~Moir, and N.~Shavit.
\newblock A dynamic-sized nonblocking work stealing deque.
\newblock \emph{Distributed Computing}, 18\penalty0 (3):\penalty0 189--207,
  2006.
\newblock \url{http://dx.doi.org/10.1007/s00446-005-0144-5}.

\bibitem[Henzinger et~al.(2013)Henzinger, Sezgin, and
  Vafeiadis]{conf/concur/HenzingerSV13}
T.~A. Henzinger, A.~Sezgin, and V.~Vafeiadis.
\newblock Aspect-oriented linearizability proofs.
\newblock In \emph{{CONCUR} 2013 - Concurrency Theory - 24th International
  Conference, {CONCUR} 2013, Buenos Aires, Argentina, August 27-30, 2013.
  Proceedings}, volume 8052 of \emph{Lecture Notes in Computer Science}, pages
  242--256. Springer, 2013.
\newblock \url{http://dx.doi.org/10.1007/978-3-642-40184-8_18}.

\bibitem[Herlihy and Wing(1990)]{journals/toplas/HerlihyW90}
M.~Herlihy and J.~M. Wing.
\newblock Linearizability: {A} correctness condition for concurrent objects.
\newblock \emph{{ACM} Trans. Program. Lang. Syst.}, 12\penalty0 (3):\penalty0
  463--492, 1990.
\newblock \url{http://doi.acm.org/10.1145/78969.78972}.

\bibitem[Liang and Feng(2013)]{conf/pldi/LiangF13}
H.~Liang and X.~Feng.
\newblock Modular verification of linearizability with non-fixed linearization
  points.
\newblock In \emph{{ACM} {SIGPLAN} Conference on Programming Language Design
  and Implementation, {PLDI} '13, Seattle, WA, USA, June 16-19, 2013}, pages
  459--470. {ACM}, 2013.
\newblock \url{http://doi.acm.org/10.1145/2462156.2462189}.

\bibitem[Liu et~al.(2009)Liu, Chen, Liu, and Sun]{conf/fm/LiuCLS09}
Y.~Liu, W.~Chen, Y.~A. Liu, and J.~Sun.
\newblock Model checking linearizability via refinement.
\newblock In \emph{{FM} 2009: Formal Methods, Second World Congress, Eindhoven,
  The Netherlands, November 2-6, 2009. Proceedings}, volume 5850 of
  \emph{Lecture Notes in Computer Science}, pages 321--337. Springer, 2009.
\newblock \url{http://dx.doi.org/10.1007/978-3-642-05089-3_21}.

\bibitem[O'Hearn et~al.(2010)O'Hearn, Rinetzky, Vechev, Yahav, and
  Yorsh]{conf/podc/OHearnRVYY10}
P.~W. O'Hearn, N.~Rinetzky, M.~T. Vechev, E.~Yahav, and G.~Yorsh.
\newblock Verifying linearizability with hindsight.
\newblock In \emph{Proceedings of the 29th Annual {ACM} Symposium on Principles
  of Distributed Computing, {PODC} 2010, Zurich, Switzerland, July 25-28,
  2010}, pages 85--94. {ACM}, 2010.
\newblock \url{http://doi.acm.org/10.1145/1835698.1835722}.

\bibitem[Shacham et~al.(2011)Shacham, Bronson, Aiken, Sagiv, Vechev, and
  Yahav]{conf/oopsla/ShachamBASVY11}
O.~Shacham, N.~G. Bronson, A.~Aiken, M.~Sagiv, M.~T. Vechev, and E.~Yahav.
\newblock Testing atomicity of composed concurrent operations.
\newblock In \emph{Proceedings of the 26th Annual {ACM} {SIGPLAN} Conference on
  Object-Oriented Programming, Systems, Languages, and Applications, {OOPSLA}
  2011, part of {SPLASH} 2011, Portland, OR, USA, October 22 - 27, 2011}, pages
  51--64. {ACM}, 2011.
\newblock \url{http://doi.acm.org/10.1145/2048066.2048073}.

\bibitem[Vafeiadis(2010)]{conf/cav/Vafeiadis10}
V.~Vafeiadis.
\newblock Automatically proving linearizability.
\newblock In \emph{Computer Aided Verification, 22nd International Conference,
  {CAV} 2010, Edinburgh, UK, July 15-19, 2010. Proceedings}, volume 6174 of
  \emph{Lecture Notes in Computer Science}, pages 450--464. Springer, 2010.
\newblock \url{http://dx.doi.org/10.1007/978-3-642-14295-6_40}.

\bibitem[Wing and Gong(1993)]{journals/jpdc/WingG93}
J.~M. Wing and C.~Gong.
\newblock Testing and verifying concurrent objects.
\newblock \emph{J. Parallel Distrib. Comput.}, 17\penalty0 (1-2):\penalty0
  164--182, 1993.
\newblock \url{http://dx.doi.org/10.1006/jpdc.1993.1015}.

\bibitem[Zhang et~al.(2013)Zhang, Chattopadhyay, and Wang]{conf/kbse/ZhangCW13}
L.~Zhang, A.~Chattopadhyay, and C.~Wang.
\newblock Round-up: Runtime checking quasi linearizability of concurrent data
  structures.
\newblock In \emph{2013 28th {IEEE/ACM} International Conference on Automated
  Software Engineering, {ASE} 2013, Silicon Valley, CA, USA, November 11-15,
  2013}, pages 4--14. {IEEE}, 2013.
\newblock \url{http://dx.doi.org/10.1109/ASE.2013.6693061}.

\bibitem[Zhang(2011)]{conf/icse/Zhang11a}
S.~J. Zhang.
\newblock Scalable automatic linearizability checking.
\newblock In \emph{Proceedings of the 33rd International Conference on Software
  Engineering, {ICSE} 2011, Waikiki, Honolulu , HI, USA, May 21-28, 2011},
  pages 1185--1187. {ACM}, 2011.
\newblock \url{http://doi.acm.org/10.1145/1985793.1986037}.

\end{thebibliography}

\end{document}